Computable Compressed Matrices\\Suplementary Material

### 2014-01-27 manuscript.tex
Figuras ordenadas corretamente.
Atualizado o texto...

Inserir os resultados

@@

An example of the application of the methodology presented is shown in Figure \ref{cf}. In this figure, the matrix with 
normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
Therefore, it becomes possible to apply the selected algorithms.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{cf}
  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
bitstring formats.}
  \label{cf}
\end{figure}

Inserir o segundo conjunto de resultados aqui.

@@

@@

>>>>

In Figures \ref{fig:23242526} and \ref{fig:27282930} the results from the processing time (in milliseconds) 
for the operations that involved only integer ($Z$) and positive integer numbers ($Z^+$) are represented. 
The traditional method (Normal) and the Bitstring method (Bitstring $Z$, integer numbers and Bitstring Zp, 
positive integer numbers) were used. The operations were attribution ($A=5$, A is a matrix with dimension n by n); 
sum of the elements of a matrix with a constant ($A=A+5$); sum of matrices ($A=A+A$); andsubtraction of the elements 
of a matrix with a constant ($A=A-8$). These operations are represented in Figure \ref{fig:23242526}. 
Figure \ref{fig:27282930}, in turn, shows the results for the multiplicationoperations of matrices ($A = A \times A$); 
for the multiplication of the the elements of a matrix by a constant ($A=A \times 2$); 
and for the calculation of transpose and maximum of a matrix.

The results show that the bitstring method is faster than the traditional ones only for the attribution.
This occurs because the element of the matrix is used more than once to store the numbers, differently 
from the traditional method. Due to the number of manipulations  the bitstring method does, the other 
operations  demand more time to be processed, but this can be optimized. When applied to positive integers, 
the bitstring model works faster than when applied to integers because it does not need to do operations 
related to the signals.

The results for the real numbers are represented in Figures \ref{fig:31323334} and \ref{fig:35363738}. 
By repeating the operations tested in the integer numbers, it was possible to see that the bitstring method 
cannot overcome the traditional method. This happensbecause its application to  the real numbers, as alternative 
to maintain the compression, asks for a conversion of real to integer numbers in order for them to be stored and 
when accessed, another conversion is made. This is a problem from the implementation process, but it can be optimized.

@@

An example of the application of the methodology presented is shown in Figure \ref{fig22}. In this figure, the matrix 
with normal and bitstring formats are displayed. In both matrices, the information is represented and can be accessed. 
Therefore, it becomes possible to apply the selected algorithms.

In order to test a possible model application, we considered a matrix with the following dimensions: 95,000 users 
(lines) by 3,000 movies (columns), with the ratings for Netflix’s movies
\footnote{http://graphlab.org/downloads/datasets/}. In order to evaluate a larger dataset, we modified the 
NetFlix data so we could increase the dimension of the matrix to be analyzed. The 95,000 users correspond to 
thelines of the matrix. In orderto gradually increase the number of lines, we performed consecutive re-samplings 
and added them to the original matrix . As a results, we obtained a  matrix with 600,000 lines and 3,000 columns. 
The operations used with the matrix were the Per User Average (PUA) and the rating estimation of a movie i using 
the Bias From Mean (BFM) \cite{cf}. The analyses were  made, considering the created matrix with different dimensions, 
with n from 40,000 to 600,000 by 20,000. It is important to emphasize that the number of columns is constant and 
equal to 3,000. The implementations of this problem were made in Fortran and both the developedlibrary based in 
the method SM and in the traditional form were used. Codes were inserted in the library in order to optimize the 
multicore processing using threads. No additional configuration was necessary to use the multicore processing. 
since the library detects the characteristics of the processor uses them in the simulations. In turn, for the 
traditional implementation, weused the optimization process and parallelization, which are made available by 
the processor,as the vectorization and multicore parallelization.

The results are represented in Figures \ref{fig39}, \ref{fig40} and \ref{fig41} for the PUA Method and in Figures 26, 27 and 28 for the 
BFM Methods.  A computer with 8 GB of RAM Memory, 16 GB of swap, processor core i7 model 870\todo{confirmar modelo}
and operational 
system Linux Debian 7.0 was used to perform this simulation. In Figure \ref{fig39}, it is possible to verify the approximate 
moment  when the  the storage in disk (swaap) begins to be used. This is indicated by the vertical green line. 
At this moment, the bitstring method becomes faster than the traditional method because it does not require disk 
storage due tocompressed memory representation. Note that, in the Figure \ref{fig40}, the traditional method is faster 
than the bitstring one before the beginning of the disk access. However, after it becomes necessary use the disk 
so that the matrix can be accessed and stored, , the traditional method  becames slower (Figura \ref{fig41}). 
These results 
indicate that the bistring method can promote the optimization of the memory use because the access to the main 
memories (RAM and cache) are faster than the access to secondary memories (disk). In Figures \ref{fig42}, 
\ref{fig43} and \ref{fig44} the 
results of the processing time for  the BFM method are represented. Once again, a vertical green line indicates 
the approximate moment when the disk access begins. There is a difference in relation to the PUA method since the 
BMF method uses the PUA values to perform its calculation. Therefore, the BMF method demands more memory to be 
processed, which reduces the matrix dimensions involved in the process, represented in the main memory. As 
indicated in Figure \ref{fig41}, for the PUA method, and Figure \ref{fig44}, for the BMF method, it is possible 
to verify that 
before the access to the disk, the traditional method is faster than the bitstring one, but once the access 
is initiated,  the bitstring method becames faster.

In spite of the simplicity of the application, it is relevant since  it allows for large-sized matrices to be 
allocated in memory and manipulated.

@@

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.4,trim=5cm 0cm 5cm 0cm,clip]{fig22}
  \caption{Illustration of the process of the bistring model application to matrices of collaborative filttering of the 
algorithms per use average (PUA) and of the bias from mean (BFM) in a matrix represented in the traditional and 
bitstring formats.}
  \label{fig22}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$A=5$]{
  \includegraphics[scale=0.4,clip]{fig23}
  }
  \subfigure[$A=A+5$]{
  \includegraphics[scale=0.4,clip]{fig24}
  }
  \subfigure[$A=A+A$]{
  \includegraphics[scale=0.4,clip]{fig25}
  }
  \subfigure[$A=A-8$]{
  \includegraphics[scale=0.4,clip]{fig26}
  }
  \caption{Processing time of the following operations: attribution ($A=5$, A is the matrix with an n by n dimension); 
  element sum of a matrix with a constant ($A=A+5$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
  matrix by a constant ($A=A-8$) involving integer numbers ($Z$) and positive integer numbers ($Z^+$). The bitstring 
  method presented the best overall performance only for the attribution operations, because the storage of 
  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
  access the elements. This can still be optimized in the implemented library.}
  \label{fig:23242526}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$A = A \times A$]{
  \includegraphics[scale=0.4,clip]{fig27}
  }
  \subfigure[$A=A \times 2$]{
  \includegraphics[scale=0.4,clip]{fig28}
  }
  \subfigure[$transpose(A)$]{
  \includegraphics[scale=0.4,clip]{fig29}
  }
  \subfigure[$maximum(A)$]{
  \includegraphics[scale=0.4,clip]{fig30}
  }
  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
  element multiplication of matrix by constant ($A=A \times 2$), the calculation of transpose and maximum of a 
  matrix involve integer numbers ($Z$) and positive integer numbers ($Z^+$). In this cases, the traditional 
  method was faster than the bitstring model.}
  \label{fig:27282930}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$A=5.0000$]{
   \includegraphics[scale=0.4,clip]{fig31}
  }
  \subfigure[$A=A+5.0000$]{
   \includegraphics[scale=0.4,clip]{fig32}
  }
  \subfigure[$A=A+A$]{
   \includegraphics[scale=0.4,clip]{fig33}
  }
  \subfigure[$A=A-8.0000$]{
   \includegraphics[scale=0.4,clip]{fig34}
  }
  \caption{Processing time of the following operations: attribution ($A=5.0000$, A is the matrix with an n by n dimension); 
  element sum of a matrix with a constant ($A=A+5.000$);sum of matrices ($A=A+A$); and subtraction of the elements of a 
  matrix by a constant ($A=A-8.0000$) involving real numbers ($R$) and positive integer numbers ($R^+$). The bitstring 
  method presented the best overall performance only for the attribution operations, because the storage of 
  different numbers occurs in one element of a matrix, which reduces the accessing time. For the other operations, 
  the bitstring method is slower due to the number ofoperations which are necessary to be  executed in order to 
  access the elements. This can still be optimized in the implemented library.}
  \label{fig:31323334}
\end{figure}

\begin{figure}[h]
  \centering
  \subfigure[$A = A \times A$]{
   \includegraphics[scale=0.4,clip]{fig35}
  }
  \subfigure[$A=A \times 2$]{
   \includegraphics[scale=0.4,clip]{fig36}
  }
  \subfigure[$transpose(A)$]{
   \includegraphics[scale=0.4,clip]{fig37}
  }
  \subfigure[$maximum(A)$]{
   \includegraphics[scale=0.4,clip]{fig38}
  }
  \caption{Processing time of matrix multiplication ($A = A \times A$, A is the matrix with a n by n dimension), 
  element multiplication of matrix by constant ($A=A \times 2.0000$), the calculation of transpose and maximum of a 
  matrix involve integer numbers ($R$) and positive integer numbers ($R^+$). In this cases, the traditional 
  method was faster than the bitstring model.}
  \label{fig:35363738}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig39}
  \caption{Processing time for the Per User Average method (PUA) in function of the numbers of linesof a matrix 
  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
  method fully represents a matrix  in memory, without the necessity to access the  disk.}
  \label{fig39}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig40}
  \caption{Processing time for the Per User Average method (PUA) in function of the  numbers of lines of a matrix 
  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
  to emphasize that the computational cost of both methods is linear.}
  \label{fig40}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig41}
  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
  method is the bitstring because of the compressed representation in the matrix memory.}
  \label{fig41}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig42}
  \caption{Processing time for the Bias From Mean method (BMF) in function of the numbers of linesof a matrix 
  with 3,000 columns. The vertical green line indicates the approximate moment when the traditional method  
  demands  disk access in order to store the matrix to be analyzed, which  overloads the process. The bitstring 
  method fully represents a matrix  in memory, without the necessity to access the  disk.}
  \label{fig42}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig43}
  \caption{Processing time for the Bias From Mean method (BMF) in function of the  numbers of lines of a matrix 
  with 3,000 columns.  The vertical green line indicates the approximate moment when the traditional method  
  demands disk access in order to store the matrixto be analyzed, which  overloads the process. For dimensions 
  inferior to approximately 400 lines, the traditional method is faster than the bitstring one. It is important 
  to emphasize that the computational cost of both methods is linear.}
  \label{fig43}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6,clip]{fig44}
  \caption{Relation between the processing time using the bitstring method and the traditional method in function 
  of the line numbers of a matrix with 3,000 columns. It is possible to verify that for a matrix with less than 400 
  lines, the traditional method is faster. However, when the number of lines is greater  than 400 lines, the best 
  method is the bitstring because of the compressed representation in the matrix memory.}
  \label{fig44}
\end{figure}

@@

\newpage

\clearpage
